### 2022-01-04 index.Rmd
Add an index file that uses jdssv format

---
documentclass: jdssv
author:
  - name: Susan Vanderplas
    affiliation: University of Nebraska, Lincoln
    # use this syntax to add text on several lines
    address: |
      | Statistics Department
      | 350 Hardin Hall North Wing
      | 3310 Holdrege St
      | University of Nebraska, Lincoln
      | Lincoln, NE 68583-0963, United States
    email: \email{svanderplas2@unl.edu}
    url: https://srvanderplas.netlify.app
  - name: Adalbert Wilhelm
    affiliation: Jacobs University
    # use this syntax to add text on several lines
    address: |
      | some address
    email: \email{a.wilhelm@jacobs-university.de}
    url: https://www.jacobs-university.de/directory/awilhelm # This link does not work...
title:
  formatted: "Covid Graphics" # We need a better title, sigh.
  # If you use tex in the formatted title, also supply version without
  plain:     "Guidelines for the Journal of Data Science, Statistics, and Visualisation Version 0.3"
  # For running headers, if needed
  short:     "Covid Graphics"
abstract: >
    Abstract text here
keywords:
  # at least one keyword must be supplied
  formatted: [COVID-19, visualization]
  plain:     [COVID-19, visualization]
preamble: >
  \usepackage{amsmath}
  \usepackage{longtable}
  \usepackage{booktabs}
  \newcommand{\class}[1]{`\code{#1}'}
  \newcommand{\fct}[1]{\code{#1()}}
  \newcommand{\ma}[1]{\ensuremath{\mathbf{#1}}}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

\newpage
# Computational Details {-}

<!-- If necessary or useful, information about certain computational details -->
<!-- such as version numbers, operating systems, or compilers could be included -->
<!-- in an unnumbered section. Also, auxiliary packages (say, for visualizations, -->
<!-- maps, tables, \dots) that are not cited in the main text can be credited here. -->

<!-- The results in this paper were obtained using -->
<!-- \proglang{R}~3.5.1. \proglang{R} itself -->
<!-- and all packages used are available from the Comprehensive -->
<!-- \proglang{R} Archive Network (CRAN) at -->
<!-- \url{https://CRAN.R-project.org/}. -->

# Acknowledgments {-}

<!-- All acknowledgments should be collected in this -->
<!-- unnumbered section before the references. It may contain the usual information -->
<!-- about funding and feedback from colleagues/reviewers/etc. Furthermore, -->
<!-- information such as relative contributions of the authors may be added here -->
<!-- (if any). -->

\bibliography{refs}

\newpage

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

# More Technical Details {#app:technical}

<!-- Appendices can be included after the bibliography (with a page break). Each section within the appendix should have a proper section title (rather than just \emph{Appendix}). -->